---
title: 'Cow Milk Analysis '
output:
 # pdf_document
  html_notebook 
 # html_document
---

#Frame
Given cow milk production from 1962-1975, explored the data in following manner.

   1. Visualize the data 
   2. visualize the seasonal trend if exists
   3. Check the stationarity of data
   4. Use ARIMA model to predict the trend for next five years
   5. Plot the forecast data to give the overview of the trend 


....


#Acquire 

Data is taken from the https://datamarket.com

It contains the monthly milk production  data info 


```{r}
df=read.csv("monthly-milk-production-pounds-p.csv",header=TRUE,stringsAsFactors=F)
str(df)
head(df)
```
Insigths:
 --Data is given monthly quantity of milk from 1962-1975
 --There is no NA or empty data
 --last two rows is not valid as it is just statement written at the end of csv file

#Refine 

*************************************************************************
        Rename the column name as it is big and NA rows from the tail
***************************************************************************

```{r}
  
colnames(df)<-c("Month","Quantity")
head(df)

#remove the last two row as it is NA 
df<-head(df,-1)
tail(df)


```

Insights
-- valid /clean data with renamed column name month and Quantity

#Explore 

*************************************************************************************
          create time series with the given data frame for further prediction of data
**************************************************************************************

```{r}

library(tseries)
library(xts)  #extended time series 


ts_milkPr <- xts(df[,-1],order.by =as.yearmon(df$Month,"%Y-%m"))

ts_milkPr  #display the data of the time series 

#plot the time series to see the trend with the mean
plot(ts_milkPr,xlab='year' ,ylab='milk quantity',main='Milk Production')
abline(a=mean(ts_milkPr),b=0, col = "red")


```


Insights

--From the above plot it is visible that the data has seasonal trending 
--seasonal trending is in additive in nature as it is not growing with level of time .
--there is no visible outliers


**************************************************************************
        Break the data into trend,seasonal variation and irregular component 
***************************************************************************

```{r}

data_decomposed <- stl(ts_milkPr, t.window=50, s.window='periodic', robust=TRUE)

data_decomposed
plot(data_decomposed)


```
Insigths:
--from the above plot can see that trend ,seasonal data ,remainder as compared to original data 
--trend is growing but at the end it becomes constant
--seaonal shows a peak in production in the mid of every year.

#Transform

***************************************************************************
        Retrieve seasonally adjusted data 
***************************************************************************

```{r}
library(forecast)

data_deseasonal=seasadj(data_decomposed)
data_deseasonal

```
Insights:
--above data is without seaonal trend component
--this will be used for further prediction


*********************************************************
       Check the stationarity of the data
*********************************************************

```{r}
#Dicky Fuller test

adf.test(ts_milkPr)

```
Insights:

--from the above result ,given time series is stationary as p-value is very small hence non stationary hypothesis is rejected and given data is stationary

****************************************************************************
          Visualize the series ACF/PACF 
******************************************************************************

```{r}
acf(ts_milkPr)
pacf(ts_milkPr)

```

from the ACF we can see that spike at 13 lag shows that there is strong seasonality
-- spike peak at 1 and 13
-- ACF plot cuts off after the first lag hence p=0 and pacf q also=0

*******************************************************************************
        reiterate the model to find the proper order AR and MA
*********************************************************************************
```{r}
data_d1 = diff(data_deseasonal, differences = 1)
plot(data_d1)  #plot the difference

adf.test(data_d1, alternative = "stationary")
```
Insights
---No visible string trend ,data is near zero


*************************************************************
                      AR/MA
*****************************************************************
```{r}
Acf(data_d1,main="ACF of D1")
Pacf(data_d1,main="Pacf of D1")

```



#Model

********************************************************
         Fit AUTO Arima Model 
********************************************************
```{r}

#auto.arima model take the AR,MA,D values itself

fit<-auto.arima(data_deseasonal,seasonal = FALSE)
fit

tsdisplay(residuals(fit), lag.max=15, main='Seasonal Model Residuals')


```

Insights:
From the above results
--Residuals plots show a smaller error range, more or less centered around 0
--AIC is smaller for the 2,1,2 range hence predict the data with this model

************************************************************
            Forecast the data for next five year
***********************************************************

```{r}
fcast<-forecast(fit,h=12*5)
plot(fcast)

```
Insights: 
--From the above model we can see the forecast in the blue line which is having the growing trend

***********************************************************
           Forecasting with seasonal effect
*************************************************************

```{r}
fit_season=auto.arima(data_deseasonal,seasonal = TRUE)
fit_season

tsdisplay(residuals(fit_season), lag.max=15, main='Seasonal Model Residuals')


```
forcast with the seasonal data

```{r}

fcast_season<-forecast(fit_season,h=12*5)
plot(fcast_season)
```

insights
: from the above forcast plot 80% confidence limits shaded in darker blue, and  95% in lighter blue.
: Forecast of data with seasonal trend is having small difference as comapred to forecast of data     without seasonal trend





