---
title: "India Health Care  Center Analysis"
output: html_notebook
---

#Frame
Given Health Care data explored in following manner.

1.Overall India Health Analysis
2.State wise presentation of the data
etc..   fill as when required


#Acquire

data is scrapped from the https://data.gov.in/

Insight :
It conatins the number of state-wise and district-wise availability of Sub-Centres, Primary Health Centres, Community Health Centres, Sub Divisional Hospitals and District hospitals In India till march 2016.

Following Abbreviations is used for the following

Primary Health Centers: PHC
Community Health Centers :CHC
Sub divisional hospitals : SDH
District Hospital : DH
SubCenters



```{r}


##### reading the data from .csv/xls files########################

df=read.csv("District-wise_Health_Centres_March_2016.csv",header=TRUE)
str(df)






```

Insigths
1. From the above table no missing value and NA Values
2. 0 value is present if corresponding health care doesnt present
3. All the columns contains the data as mentioned in the data overview
4. column name is too long and needs to be renamed for the further analysis



#Refine 

Step1:
*************************************************************
Rename the column name as it is very long
************************************************************

```{r}

newcolnames<-c('S.No.','states','District','#SubCentres','#PHC','#CHC','#SDH','#DH')
colnames(df)<-newcolnames
head(df)



```

Insights:

From the above data , needs to change the data type of the first three column as it is factor 


Step2:
****************************************************************
Change the Data Type of the column
****************************************************************

```{r}


df$S.No.<- as.numeric(as.character(df$S.No.))
df$states<-as.character(df$states)
df$District<-as.character(df$District)


#check the kind of data it is storing  
head(df,50)
tail(df,50)


```

Insights

1. Each states contain district-wise data as well as contain the total state wise  data also
2. last row contains the all india total data 

......Add as and when required 


#Transform

.....needs to be done 


#Explore 
1. Visualise the All India  data in the pie chart to give the overall idea of the different type of health centers 

```{r}

library(dplyr)
library(ggplot2)
#retrive the last row of data frame which contain the all india health centers
tail(df,1,header=TRUE)


dc<- tail(df,1)
dat <- as.data.frame(t(dc))
      


#take only last  five rows of the dataframe
dat <-tail(dat,5)
dat
dat$type<-c('#SubCenters','#PHC','#CHC','#SDH','#DH')

dat <-dat[,c(2,1)]

#assign the column names 
colnames(dat)<-c("HealthCenterType","TotalCount")

#convert the column data type as integer from factor
dat$TotalCount=as.numeric(as.character(dat$TotalCount))

dat

#Pie Chart
dat$frac=dat$TotalCount/sum(dat$TotalCount)

dat
dat = dat[order(dat$frac),]

dat$ymax = cumsum(dat$frac)
dat$ymin = c(0, head(dat$ymax, n=-1))

#p2 = ggplot(dat, aes(fill=HealthCenterType, ymax=ymax, ymin=ymin, xmax=4,min=3))+
 #    geom_rect(colour="grey30") +
  #   coord_polar(theta="y") +
   #  xlim(c(0, 4)) 
 #    theme_bw() +
 #    theme(panel.grid=element_blank()) +
 #    theme(axis.text=element_blank()) +
 #    theme(axis.ticks=element_blank()) +
  #   labs(title="Customized ring plot")

dchart=ggplot(dat) + geom_rect(aes(fill=HealthCenterType, ymax=ymax, ymin=ymin, xmax=4, xmin=3)) +
coord_polar(theta="y") + xlim(c(1, 4))


dchart+  geom_label(aes(label=paste(as.integer(frac*100),"%"),x=3.5,y=(ymin+ymax)/2),inherit.aes = TRUE, show.legend = FALSE)

dat
```


Insigths
From the above we can see that in india subcenter are maximum and District hospital are least 
..to be done the label properly on donut chart  as label are not 


Step2:
1.find if there is relation between number size of the stat and the number of total health care center
2. More number of district depicts the size of the state.
3. Add one more column for  number of district
4. One more column for the total number of health center in particular state

```{r}

library(dplyr)

#select only s.no states ,#SubCenters,#PHC,#CHC,#SDH,#DH and create a new data frame FOR statewise report 
df_StateWiseData <-df[,c("states","#SubCentres","#PHC","#CHC","#SDH","#DH")]

#Add the new column district having value 1 to get the number of district 
df_StateWiseData$District <- rep(1,nrow(df_StateWiseData))

df_StateWiseData <- df_StateWiseData %>%
                    group_by(states) %>%
                    summarise_each(funs(sum))
                      
                  
                     



                    

  df_StateWiseData <- df_StateWiseData %>%
                      filter(states!="All India/ Total")
  
  #reorder the column by changing the index 

df_StateWiseData<-  df_StateWiseData[,c(1,7,2,3,4,5,6)]

head(df_StateWiseData,20)


```


Step2: Find the total number of Health center in a states ,add one extra column which will tell the total number of states.


```{r}

df_StateWiseData$total_HealthCenter <- df_StateWiseData$`#SubCentres` +
                                       df_StateWiseData$`#PHC`   +
                                       df_StateWiseData$`#CHC` +
                                       df_StateWiseData$`#SDH` +
                                       df_StateWiseData$`#DH`

head(df_StateWiseData,20)
         
  
```

Step3:Find the relation if number of health center is depending on the number of district
1. Draw the scatter plot  

```{r}

library(ggplot2)
#data for regression  district and totalHealthCenter

df_reg <-df_StateWiseData[,c(2,8)]

#attach(df_StateWiseData)
#df_StateWiseData %>%
#                 plot(total_HealthCenter,District, main = "HealthCenter in India",xlab = #"Number of                     District",ylab = "total_HealthCenter",
#
df_reg
df_reg <-df_reg[order(District),]

df_reg


#scatter diagram

plot(df_reg$District,df_reg$total_HealthCenter,xlab="District",ylab="Health centers")
abline(lm(df_reg$total_HealthCenter ~ df_reg$District)) 



```

Create the realtionship model between Number of district and health center opened
--Assuming
H0:b1=0 ,there is no relationship between the number of district and health care center
H1:b1!=0 ,there is relatiom ship between the two variables

```{r}
relation <- lm(formula =total_HealthCenter ~ District,data=df_reg)

print(summary(relation))


```

Insights:
Residuals are the difference between observed and the response  value that is predicted  from the model.

2. here from the result three significant star "***" ,means we reject the Null hypothesis that there is no relationship between size of the state and the Health center opened.

3.Intercept Estimation : when district is zero then average value of health center is -2509.85
4.slope estimated value :given one unit change in district expected change in healthcare is 661.65

5.Standard error for the slope is :59.33 which menas here variablity in coefficient value across the sample ,it is almost 11 times lesser than the coefficient value

6.probablityof coefficient is nearly zero and R2 is .78

7.From all the above point we reject the null hypothesis ,Number of health care opened is ~78 dependent on the size of state



*******************************************************************************

predict the number of health care opened when the number of district  in a state is 24 and 25


```{r}
new_df <-data.frame(District=c(24,25))

total_hc <- predict(relation,new_df)

print(total_hc)


```

****************************************************************************************
star plot for statewise Data

```{r}
require(grDevices)
data(df_StateWiseData)
palette(rainbow(12, s = 0.6, v = 0.75))

stars(df_StateWiseData[, 1:7], key.loc = c(14, 2),
      main = "HealthCenters : stars(*, full = F)", full = FALSE)

stars(df_StateWiseData[, 1:7], key.loc = c(14, 1.5),col.segments = 1:nrow(df_StateWiseData) 
      ,draw.segments = TRUE,
      main = "HealthCenters
      : full stars()", flip.labels = TRUE)
```

