---
title: "Oniona DateWise Analysis "
output: html_notebook
---


#Frame
Question to be answered
1.which city has the highest sale of onion in a year?
2.predict the price of onion for the next 30 days?


#Acquire
As per question scrape data from NHRFD on Datewise basis
taken 2016 datewise data 
--converted the date into .csv format and then reading it in dataframe


```{r}
############# reading the data from html file##############################
#library(rvest)
#Read the data from the file 
#pg.out<-read_html("DailyWiseMarketArrivals.htm")
#pg.out %>%
 #    html_node('#dnn_ctr966_DailyWiseMarketArrivals_GridView1') %>%
  #   html_table()
# df <-pg.table
############################################################################

########################## reading the data from .csv/xls files########################

df=read.csv("DailyWiseMarketArrivals.csv",header=TRUE)
str(df)






```

#Refine
Refine the received raw data into fine data for the analysis

Step1:
*****************************************************************************
-Rename the column name as it is long so assigning new name to column
*****************************************************************************
```{r}
newnames<-c('date','market','quantity','pricemin','pricemax','pricemod')
colnames(df)<-newnames
head(df)

```

Step2:
*****************************************************************
-Remove the last row as it contains the total of all the  data 
******************************************************************


```{r}
library(dplyr)

df<-df %>%
  filter(market!='Total') 
tail(df)


```

Step3:
**************************************************************
-change the data type of the columns
**************************************************************
```{r}

library(dplyr)

df$date<-as.Date(df$date, format = "%d/%b/%Y")  #date format date-month-year
df$market<-as.character(df$market)
df$pricemin<-as.numeric(as.character(df$pricemin))
df$pricemax<-as.numeric(as.character(df$pricemax))
df$pricemod<-as.numeric(as.character(df$pricemod))

str(df)



```
step4:

*************************************************************
--Retrive  the state name as city and state name is combined
*************************************************************
```{r}

library(tidyr)

str(df)

unique(df$market)
df<-df %>%
    mutate(market1=market )%>%
    separate(market1,c('city','state'),sep="\\(")

#remove the last ")" from the state
df$state<-substr(df$state,1,nchar(df$state)-1) #removing thelast ) from state

str(df)
```
step5:
*********************************************************************************
--update the state code where state code is not mentioned hardcoding the value
*********************************************************************************

```{r}
head(df)

df$state=ifelse(df$city=='BANGALORE','KNT',df$state)
df$state=ifelse(df$city=='CHENNAI','TN',df$state)
df$state=ifelse(df$city=='DELHI','DL',df$state)
df$state=ifelse(df$city=='BHOPAL','MP',df$state)
df$state=ifelse(df$city=='CHANDIGARH','CH',df$state)
df$state=ifelse(df$city=='GUWAHATI','AS',df$state)
df$state=ifelse(df$city=='HYDERABAD','TELANGANA',df$state)
df$state=ifelse(df$city=='JAIPUR','RAJ',df$state)
df$state=ifelse(df$city=='JAMMU','JK',df$state)
df$state=ifelse(df$city=='KOLKATA','WB',df$state)
df$state=ifelse(df$city=='LUCKNOW','UP',df$state)
df$state=ifelse(df$city=='MUMBAI','MS',df$state)
df$state=ifelse(df$city=='NAGPUR','MS',df$state)
df$state=ifelse(df$city=='PATNA','BHR',df$state)
df$state=ifelse(df$city=='SHAHJAHANPUR','UP',df$state)
df$state=ifelse(df$city=='TIPHRA','CG',df$state)
df$state=ifelse(df$city=='BANGALORE','KNT',df$state)

   

```


Step6:
***************************************************************************************
--function to convert the state code into the state name and add the column to dataframe
****************************************************************************************

```{r}

stateFromStateCode <-function(x)
{
  st.codes<-data.frame(
  
    state=as.factor(c("AN", "AP", "AR", "AS", "BHR", "CH", "CG", "DN", "DD", "DL",    "GA","GUJ","HR","HP","JK","JH","KNT","KER","LD","MP","MS","MN","ML","MZ","NL","OD","PY","PB","RAJ","SK","TN","TELANGANA","TR","UK","UP","WB")),
    
    fullName=as.factor(c("Andaman and Nicobar Islands","Andhra Pradesh","Arunachal Pradesh","Assam","Bihar","Chandigarh","Chhattisgarh","Dadra and Nagar Haveli","Daman and Diu","Delhi","Goa","Gujrat","Haryana","Himachal Pradesh","Jammu and Kashmir","Jharkhand","Karnataka","Kerala","Lakshadweep","Madhya Pradesh","Maharashtra","Manipur","Meghalaya","Mizoram","Nagaland","Odisha","Puducherry","Punjab","Rajasthan","Sikkim","Tamil Nadu","Telangana","Tripura","Uttarakhand","Uttar Pradesh","West Bengal"))
    
                         )
  
   #create an  data.frame of state codes from source column
    st.x<-data.frame(state=x)
    #match source codes with codes from 'st.codes' local variable and use to return the full state name
    refac.x<-st.codes$full[match(st.x$state,st.codes$state)]
    #return the full state names in the same order in which they appeared in the original source
    return(refac.x)
   
}
df$StateName<-stateFromStateCode(df$state)

```



#Transform
*****************************************************************
--find the maximum sales quantity by grouping the data together based on state
****************************************************************
```{r}
# change the data type of statename

df$StateName<-as.character(df$StateName)
# sum the quantity of state wise

dfState <-df %>%
  group_by(StateName) %>%
  summarise(quantity_state=sum(quantity)) %>%
  arrange(desc(quantity_state)) 
  
#taking only first 6 date  

dfState <- head(dfState)



```

#Explore
*******************************************************************
--Find the satate having the maximum sales quantity by histograph
*******************************************************************

```{r}
library(ggplot2)
statePlot<-ggplot(dfState)+
aes(reorder(StateName,quantity_state),weight=quantity_state) +
geom_bar()+
  coord_flip()

library(plotly)
ggplotly(statePlot)


```
step2:
*************************************************************
--show the value in the map
*************************************************************


```{r}
library(ggmap)
#geocode("KNT
#dfState<- dfState %>%
 #         filter(state!="NA")
geo<-geocode(dfState$StateName)
str(geo)
dfStateGeo<-bind_cols(dfState,geo)
ggplot(dfStateGeo) +
  aes(lon,lat,size=quantity_state,color=StateName) +
  geom_point()+
 coord_map()
map<-get_map("India",maptype = "watercolor",source = "stamen",zoom=5)
ggmap(map)

ggmap(map)+geom_point(data=dfStateGeo,aes(lon,lat,size=quantity_state/1000,color=StateName))


```

#transform 2
****************************************************************
---since MS(Maharastra data is maximum one,so predict the price)
*****************************************************************

```{r}
dim(df)

dfMaharastra<-df %>% 
  filter(StateName =="Maharashtra")
#ggplot(dfMaharastra)+aes(date,pricemin,color=state)+geom_line()

dfTall <- dfMaharastra %>%
        gather("priceType","priceValue",4:6)

head(dfTall)
ggplot(dfTall) +
  aes(date,priceValue,color=priceType) +geom_line()

#head(dfTall)

```

#Model 
*****************************************************
---forcast the price max for the next 30 days  
********************************************************

```{r}

ts <- df %>%
    filter (StateName=="Maharashtra") %>%
    select(date,pricemax)


#taking first 200 data otherwise data will be more
ts<-head(ts,200)
colnames(ts)<-c('ds','y')

library(prophet)
library(Rcpp)
m <- prophet(ts)

future<-make_future_dataframe(m,periods = 30)
forcast <-predict(m,future) 
head(forcast)
plot(m,forcast)


```

